---
title: "ABM_Rmarkdown"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("G:/My Drive/University of Seoul/Agent-based model")
library(ggplot2)
library(reshape2)
library(dplyr)
library(lubridate)
library(tidyr)
library(writexl)
source("inflow_history.R")
source("Agent_based_Geum_diversion_v4.R")
```

## Agent-based model

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Executing ABM

```{r include=FALSE}

set.seed(1125)

with_ABM_new_diversion <- Agent_Geum_Network_Diversion(ABM=TRUE,opt_year = 39)
without_ABM_new_diversion <- Agent_Geum_Network_Diversion(ABM=FALSE, opt_year = 39)
```

```{r}
with_ABM_new_diversion$Gini_coefficient
without_ABM_new_diversion$Gini_coefficient

```

## Including Plots

You can also embed plots, for example:

```{r  echo=FALSE}


without_ABM_new_diversion$gg_agr
ggsave("figures/fig3.png")

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r colormap, fig.height=4, fig.width=3}

# --- column labels 먼저 정의! ---
col_labels <- c("Intake", "Agr", "Power")

rows <- c(with_ABM_new_diversion$breaker$first_row,
          with_ABM_new_diversion$breaker$second_row,
          with_ABM_new_diversion$breaker$third_row)

cols <- c(with_ABM_new_diversion$breaker$first_col,
          with_ABM_new_diversion$breaker$second_col,
          with_ABM_new_diversion$breaker$third_col)

# --- NA 제거 + 범위 밖 값 제거(중요) ---
valid <- !is.na(rows) & !is.na(cols) &
         rows >= 1 & rows <= 14 &
         cols >= 1 & cols <= 3

rows <- rows[valid]
cols <- cols[valid]

# --- count matrix ---
heat_matrix <- matrix(0, nrow = 14, ncol = 3)

for (k in seq_along(rows)) {
  r <- rows[k]
  c <- cols[k]
  heat_matrix[r, c] <- heat_matrix[r, c] + 1
}

df_heat <- melt(heat_matrix)
colnames(df_heat) <- c("row", "col", "count")

# --- labels / factors for plotting ---
df_heat <- df_heat %>%
  mutate(
    row_f = factor(row, levels = 14:1),                 # show 14 at top
    col_f = factor(col, levels = 1:3, labels = col_labels),
    label = ifelse(count == 0, "", as.character(count)),
    count_plot = ifelse(count == 0, NA, count)          # optional: hide 0 tiles
  )

ggplot(df_heat, aes(x = col_f, y = row_f, fill = count_plot)) +
  geom_tile(color = "black", linewidth = 0.3) +
  geom_text(aes(label = label), size = 3) +
  scale_x_discrete(position = "top") +
  scale_y_discrete(breaks = factor(14:1), labels = 14:1) +
  scale_fill_gradient(
    low  = "white",
    high = "red",
    name = "Number of\nABM triggers",
    na.value = "white"
  ) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 12) +
  theme(
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    axis.text.x.top = element_text(size = 12),
    axis.text.y = element_text(size = 10),
    legend.position = "right"
  )

ggsave("figures/fig6.png")
```

# Cumulative plots

```{r}
DF <- data.frame(Intake = 1:12, 
                 with_ABM = with_ABM_new_diversion$Table_intake[,2], 
                 without_ABM =  without_ABM_new_diversion$Table_intake[,2])
DFlong <- DF |> pivot_longer(cols = -Intake,names_to = "ABM")

ggplot(DFlong,aes(x=Intake, y = value,fill= ABM)) + 
  geom_col(position="dodge") +
  theme_bw() +ylab("Days of Water Supply Availability")+
  ggtitle("취수장의 용수공급가능일수") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = NULL) +
  coord_cartesian(ylim = c(1500, NA))
```

```{r}

diff <- with_ABM_new_diversion$Table_agr[,2] - without_ABM_new_diversion$Table_agr[,2]

DF <- data.frame(Agr = 1:14, 
                 with_ABM = with_ABM_new_diversion$Table_agr[,2], 
                 without_ABM =  without_ABM_new_diversion$Table_agr[,2])
DFlong <- DF |> pivot_longer(cols = -Agr,names_to = "ABM")

ggplot(DFlong, aes(x = factor(Agr), y = value, fill = ABM)) + 
  geom_col(position = "dodge") +
  theme_bw() +
  ylab("Days of Water Supply Availability") +
  ggtitle("Cumulative Reliability") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05)), limits = NULL) +
  coord_cartesian(ylim = c(500, NA)) +
  scale_x_discrete(breaks = as.character(1:14))


```

```{r}
DF <- data.frame(
  Agr = 1:14, 
  with_ABM = with_ABM_new_diversion$Table_agr[,2], 
  without_ABM = without_ABM_new_diversion$Table_agr[,2]
)

DFlong <- DF |>
  tidyr::pivot_longer(cols = -Agr, names_to = "ABM")

diff <- DF$with_ABM - DF$without_ABM

df_diff <- data.frame(
  Agr  = 1:14,
  diff = diff
) |>
  dplyr::filter(!is.na(diff) & diff != 0)

ggplot(DFlong, aes(x = factor(Agr), y = value, fill = ABM)) + 
  geom_col(position = position_dodge(width = 0.9)) +
  
  geom_text(
    data = df_diff,
    aes(
      x = factor(Agr),
      y = pmax(
            DF$with_ABM[Agr],
            DF$without_ABM[Agr]
          ) + 200,
      label = sprintf("%+.1f", diff),
      color = diff > 0
    ),
    inherit.aes = FALSE,
    size = 5,
    fontface = "bold"
  ) +
  
  scale_color_manual(
    values = c("TRUE" = "darkgreen", "FALSE" = "red")
  ) +
  guides(color = "none") +
  
  scale_fill_discrete(labels = c("with_ABM" = "with ABM", 
                                 "without_ABM" = "without ABM")) +
  scale_x_discrete(
  breaks = as.character(1:14),
  labels = paste(1:14))+
  xlab("Agricultural Stakeholders")+
  theme_bw() +
  ylab("Days of Water Supply Availability") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  coord_cartesian(ylim = c(500, NA)) +
  scale_x_discrete(breaks = as.character(1:14))

ggsave("figures/fig7.png")

```



```{r}


DF <- data.frame(Power = 1:3, 
                 with_ABM = with_ABM_new_diversion$Table_power[,2], 
                 without_ABM =  without_ABM_new_diversion$Table_power[,2])
DFlong <- DF |> pivot_longer(cols = -Power,names_to = "ABM")

ggplot(DFlong,aes(x=Power, y = value,fill= ABM)) + 
  geom_col(position="dodge") +
  theme_bw() +ylab("Days of Water Supply Availability")
```

# Drought Events

```{r}
# Drought Event 1: 1994 - 1997 ------------------------------------------------------------------

ggplot(out$series, aes(x = Time, y = icdefit)) +
  geom_line(color = "red") +
  geom_hline(aes(yintercept = 0), linetype = "dashed") +
  
  # 🔶 1994–1997 강조 박스
  annotate("rect",
           xmin = as.Date("1994-01-01"),
           xmax = as.Date("1997-12-31"),
           ymin = -Inf,
           ymax = Inf,
           alpha = 0.5,          # 투명도
           fill = "grey") +
  
  # 🔶 2014–2017 강조 박스
  annotate("rect",
           xmin = as.Date("2014-01-01"),
           xmax = as.Date("2017-12-31"),
           ymin = -Inf,
           ymax = Inf,
           alpha = 0.5,
           fill = "grey") +
  
  labs(#title = "Drought Deficit Over Time",
       y = "Deficit (MCM)",
       x = "Year") +
  
  scale_x_date(
    date_breaks = "1 year",
    date_labels = "%Y"
  ) +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5)
  )


```

```{r}
library(dplyr)

# Event 구간 정의
event_windows <- tibble(
  event = c("Event 1", "Event 2"),
  start = as.Date(c("1994-01-01","2014-01-01")),
  end   = as.Date(c("1997-12-31","2017-12-31"))
)

# 구간별 duration/magnitude 계산 (out$series는 deficit 있음)
event_stats <- event_windows %>%
  rowwise() %>%
  mutate(
    duration = sum(out$series$Time >= start & out$series$Time <= end & out$series$deficit > 0),
    magnitude = sum(out$series$deficit[out$series$Time >= start & out$series$Time <= end], na.rm=TRUE),
    intensity = magnitude / duration
  ) %>%
  ungroup()

print(event_stats)

```



```{r}
ggplot(out$series, aes(x = Time, y = deficit)) +
  geom_line(color = "red") +
  geom_hline(aes(yintercept = 0), linetype = "dashed") +
  
  # 🔶 1994–1997 강조 박스 (Event 1)
  annotate("rect",
           xmin = as.Date("1994-01-01"),
           xmax = as.Date("1997-12-31"),
           ymin = -Inf,
           ymax = Inf,
           alpha = 0.5,
           fill = "grey") +
  
  annotate("text",
           x = as.Date("1996-1-15"),   # 박스 가운데쯤
           y = 590,
           label = "Event 1",
           size = 4,
           fontface = "bold") +
  
  # 🔶 2014–2017 강조 박스 (Event 2)
  annotate("rect",
           xmin = as.Date("2014-01-01"),
           xmax = as.Date("2017-12-31"),
           ymin = -Inf,
           ymax = Inf,
           alpha = 0.5,
           fill = "grey") +
  
  annotate("text",
           x = as.Date("2016-1-15"),
           y = 590,
           label = "Event 2",
           size = 4,
           fontface = "bold") +
  
  labs(
       y = "Deficit (MCM)",
       x = "Year") +
  
  scale_x_date(
    date_breaks = "1 year",
    date_labels = "%Y"
  ) +
  scale_y_continuous(limits = c(NA, 600)) +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5)
  )
ggsave("figures/fig9.png")
```



# Result

```{r}
start_date <- as.Date("1981-10-01")
n_days <- nrow(with_ABM_new_diversion$df_intake)

date_seq <- seq(start_date, by = "day", length.out = n_days)
```

# Agr 5

```{r}
ts_df <- data.frame(
  Date = date_seq,
  Intake4_withABM    = with_ABM_new_diversion$df_intake[, 4],
  Intake4_withoutABM = without_ABM_new_diversion$df_intake[,4],
  Agr5_withABM       = with_ABM_new_diversion$df_agr[, 5],
  Agr5_withoutABM    = without_ABM_new_diversion$df_agr[,5],
  Intake5_withABM       = with_ABM_new_diversion$df_intake[,5],
  Intake5_withoutABM    = without_ABM_new_diversion$df_intake[,5]
)

# long-format으로 변환 (Stakeholder × Scenario)
ts_long <- ts_df |>
  pivot_longer(
    cols = -Date,
    names_to   = c("Stakeholder", "Scenario"),
    names_pattern = "(.*)_(withABM|withoutABM)",
    values_to  = "value"
  ) |>
  mutate(
    Stakeholder = factor(Stakeholder,
                         levels = c("Intake4", "Agr5", "Intake5"),
                         labels = c("Intake 4 (tributary donor)", "Agr 5 (recipient)", "Intake 5 (main donor)")),
    Scenario = factor(Scenario,
                      levels = c("withoutABM", "withABM"),
                      labels = c("Without ABM", "With ABM"))
  )




## ===== 1. Drought Event 1 (1994–1997) =====

drought1_long <- ts_long |>
  filter(
    Date >= as.Date("1994-01-01"),
    Date <= as.Date("1997-12-31")
  )


gg_drought1_Agr5 <- ggplot(drought1_long,
                          aes(x = Date, y = value, color = Scenario, linetype = Scenario)) +
  geom_line(linewidth = 0.5) +
  #scale_y_continuous(limits = c(0.002, NA)) +
  facet_wrap(~ Stakeholder, ncol = 1, scales = "free_y") +
  
  scale_color_manual(values = c("Without ABM" = "grey40",
                                "With ABM"    = "red")) +
  labs(
    x = "Time",
    y = "Daily Supplied Water (MCM)"
  ) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold"),
    legend.position = "top",
    axis.text.x = element_text( vjust = 0.5),
    plot.title = element_text(hjust = 0.5)
  )

gg_drought1_Agr5

ggsave("figures/fig10.png")

```

```{r}
drought2_long <- ts_long |>
  filter(
    Date >= as.Date("2014-01-01"),
    Date <= as.Date("2017-12-31")
  )

gg_drought2_Agr5 <- ggplot(drought2_long,
                          aes(x = Date, y = value, color = Scenario, linetype = Scenario)) +
  geom_line(linewidth = 0.5) +
  facet_wrap(~ Stakeholder, ncol = 1, scales = "free_y") +
  scale_color_manual(values = c("Without ABM" = "grey40",
                                "With ABM"    = "red")) +
  labs(
    x = "Time",
    y = "Daily Supplied Water (MCM)"
  ) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold"),
    legend.position = "top",
    axis.text.x = element_text( vjust = 0.5),
    plot.title = element_text(hjust = 0.5)
  )

gg_drought2_Agr5

ggsave("figures/fig11.png")

```

# Agr 11

```{r}

ts_df <- data.frame(
  Date = date_seq,
  Intake11_withABM    = with_ABM_new_diversion$df_intake[, 11],
  Intake11_withoutABM = without_ABM_new_diversion$df_intake[, 11],
  Agr11_withABM       = with_ABM_new_diversion$df_agr[, 11],
  Agr11_withoutABM    = without_ABM_new_diversion$df_agr[, 11],
  Agr12_withABM       = with_ABM_new_diversion$df_agr[, 12],
  Agr12_withoutABM    = without_ABM_new_diversion$df_agr[, 12]
)

# long-format으로 변환 (Stakeholder × Scenario)
ts_long <- ts_df |>
  pivot_longer(
    cols = -Date,
    names_to   = c("Stakeholder", "Scenario"),
    names_pattern = "(.*)_(withABM|withoutABM)",
    values_to  = "value"
  ) |>
  mutate(
    Stakeholder = factor(Stakeholder,
                         levels = c("Intake11", "Agr11", "Agr12"),
                         labels = c("Intake 11 (tributary donor)", "Agr 11 (recipient)", "Agr 12 (main donor)")),
    Scenario = factor(Scenario,
                      levels = c("withoutABM", "withABM"),
                      labels = c("Without ABM", "With ABM"))
  )




## ===== 1. Drought Event 1 (1994–1997) =====

drought1_long <- ts_long |>
  filter(
    Date >= as.Date("1994-01-01"),
    Date <= as.Date("1997-12-31")
  )

gg_drought1_Agr11 <- ggplot(drought1_long,
                          aes(x = Date, y = value, color = Scenario, linetype = Scenario)) +
  geom_line(linewidth = 0.5) +
  facet_wrap(~ Stakeholder, ncol = 1, scales = "free_y") +
  scale_color_manual(values = c("Without ABM" = "grey40",
                                "With ABM"    = "red")) +
  labs(
    x = "Time",
    y = "Daily Supplied Water (MCM)"
  ) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold"),
    legend.position = "top",
    axis.text.x = element_text( vjust = 0.5),
    plot.title = element_text(hjust = 0.5)
  )

gg_drought1_Agr11

ggsave("figures/fig14.png")

```

```{r}

## ===== 2. Drought Event 2 (2014–2017) =====

drought2_long <- ts_long |>
  filter(
    Date >= as.Date("2014-01-01"),
    Date <= as.Date("2017-12-31")
  )

gg_drought2_Agr11 <- ggplot(drought2_long,
                          aes(x = Date, y = value, color = Scenario, linetype = Scenario)) +
  geom_line(linewidth = 0.5) +
  facet_wrap(~ Stakeholder, ncol = 1, scales = "free_y") +
  scale_color_manual(values = c("Without ABM" = "grey40",
                                "With ABM"    = "red")) +
  labs(
    x = "Time",
    y = "Daily Supplied Water (MCM)"
  ) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold"),
    legend.position = "top",
    axis.text.x = element_text(angle = 90, vjust = 0.5),
    plot.title = element_text(hjust = 0.5)
  )

gg_drought2_Agr11

ggsave("figures/fig15.png")

```

# Result: Agr 13

```{r}
ts_df <- data.frame(
  Date = date_seq,
  Agr13_withABM       = with_ABM_new_diversion$df_agr[, 13],
  Agr13_withoutABM    = without_ABM_new_diversion$df_agr[, 13],
  Agr14_withABM       = with_ABM_new_diversion$df_agr[, 14],
  Agr14_withoutABM    = without_ABM_new_diversion$df_agr[, 14]
)


# long-format으로 변환 (Stakeholder × Scenario)
ts_long <- ts_df |>
  pivot_longer(
    cols = -Date,
    names_to   = c("Stakeholder", "Scenario"),
    names_pattern = "(.*)_(withABM|withoutABM)",
    values_to  = "value"
  ) |>
  mutate(
    Stakeholder = factor(Stakeholder,
                         levels = c("Agr13", "Agr14"),
                         labels = c("Agr 13 (recipient)", "Agr 14 (main donor)")),
    Scenario = factor(Scenario,
                      levels = c("withoutABM", "withABM"),
                      labels = c("Without ABM", "With ABM"))
  )




## ===== 1. Drought Event 1 (1994–1997) =====

drought1_long <- ts_long |>
  filter(
    Date >= as.Date("1994-01-01"),
    Date <= as.Date("1997-12-31")
  )

gg_drought1_Agr14 <- ggplot(drought1_long,
                          aes(x = Date, y = value, color = Scenario, linetype = Scenario)) +
  geom_line(linewidth = 0.5) +
  facet_wrap(~ Stakeholder, ncol = 1, scales = "free_y") +
  scale_color_manual(values = c("Without ABM" = "grey40",
                                "With ABM"    = "red")) +
  labs(
    x = "Time",
    y = "Daily Supplied Water (MCM)"
  ) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold"),
    legend.position = "top",
    axis.text.x = element_text(vjust = 0.5),
    plot.title = element_text(hjust = 0.5)
  )

gg_drought1_Agr14

ggsave("figures/fig16.png")

```

```{r}


## ===== 2. Drought Event 2 (2014–2017) =====

drought2_long <- ts_long |>
  filter(
    Date >= as.Date("2014-01-01"),
    Date <= as.Date("2017-12-31")
  )

gg_drought2_Agr14<- ggplot(drought2_long,
                          aes(x = Date, y = value, color = Scenario, linetype = Scenario)) +
  geom_line(linewidth = 0.5) +
  facet_wrap(~ Stakeholder, ncol = 1, scales = "free_y") +
  scale_color_manual(values = c("Without ABM" = "grey40",
                                "With ABM"    = "red")) +
  labs(
    x = "Time",
    y = "Daily Supplied Water (MCM)"
  ) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold"),
    legend.position = "top",
    axis.text.x = element_text(angle = 90, vjust = 0.5),
    plot.title = element_text(hjust = 0.5)
  )


gg_drought2_Agr14

ggsave("figures/fig17.png")
```

# Result: Agr 7

```{r}

ts_df <- data.frame(
  Date = date_seq,
  Intake6_withABM    = with_ABM_new_diversion$df_intake[, 6],
  Intake6_withoutABM = without_ABM_new_diversion$df_intake[, 6],
  Agr7_withABM       = with_ABM_new_diversion$df_agr[, 7],
  Agr7_withoutABM    = without_ABM_new_diversion$df_agr[, 7],
  Intake7_withABM       = with_ABM_new_diversion$df_intake[, 7],
  Intake7_withoutABM    = without_ABM_new_diversion$df_intake[, 7]
)


# long-format으로 변환 (Stakeholder × Scenario)
ts_long <- ts_df |>
  pivot_longer(
    cols = -Date,
    names_to   = c("Stakeholder", "Scenario"),
    names_pattern = "(.*)_(withABM|withoutABM)",
    values_to  = "value"
  ) |>
  mutate(
    Stakeholder = factor(Stakeholder,
                         levels = c("Intake6", "Agr7", "Intake7"),
                         labels = c("Intake 6 (tributary donor)", "Agr 7 (recipient)", "Intake 7 (main donor)")),
    Scenario = factor(Scenario,
                      levels = c("withoutABM", "withABM"),
                      labels = c("Without ABM", "With ABM"))
  )




## ===== 1. Drought Event 1 (1994–1997) =====

drought1_long <- ts_long |>
  filter(
    Date >= as.Date("1994-01-01"),
    Date <= as.Date("1997-12-31")
  )

gg_drought1_Agr7 <- ggplot(drought1_long,
                          aes(x = Date, y = value, color = Scenario, linetype = Scenario)) +
  geom_line(linewidth = 0.5) +
  facet_wrap(~ Stakeholder, ncol = 1, scales = "free_y") +
  scale_color_manual(values = c("Without ABM" = "grey40",
                                "With ABM"    = "red")) +
  labs(
    x = "Time",
    y = "Daily Supplied Water (MCM)"
  ) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold"),
    legend.position = "top",
    axis.text.x = element_text(vjust = 0.5),
    plot.title = element_text(hjust = 0.5)
  )

gg_drought1_Agr7

ggsave("figures/fig12.png")
```

```{r}

## ===== 2. Drought Event 2 (2014–2017) =====

drought2_long <- ts_long |>
  filter(
    Date >= as.Date("2014-01-01"),
    Date <= as.Date("2017-12-31")
  )

gg_drought2_Agr7 <- ggplot(drought2_long,
                          aes(x = Date, y = value, color = Scenario, linetype = Scenario)) +
  geom_line(linewidth = 0.5) +
  facet_wrap(~ Stakeholder, ncol = 1, scales = "free_y") +
  scale_color_manual(values = c("Without ABM" = "grey40",
                                "With ABM"    = "red")) +
  labs(
    x = "Time",
    y = "Daily Supplied Water (MCM)"
  ) +
  theme_bw() +
  theme(
    #text = element_text(family = "Helvetica"),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(face = "bold"),
    legend.position = "top",
    axis.text.x = element_text(vjust = 0.5),
    plot.title = element_text(hjust = 0.5)
  )

gg_drought2_Agr7

ggsave("figures/fig13.png")

```

## Drought Event 1 (1994–1997)

```{r, fig.width=10, fig.height=5}
library(gridExtra)
library(cowplot) 
# Add a) b) labels and remove individual legends
p1 <- gg_drought1_Agr11 + 
  ggtitle("a)") + 
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0, face = "bold", size = 14))

p2 <- gg_drought1_Agr14 + 
  ggtitle("b)") + 
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0, face = "bold", size = 14),
        axis.title.y = element_blank())  # ← 오른쪽 plot y축 레이블 제거

main_grid <- plot_grid(p1, p2, ncol = 2, align = "hv", axis = "tblr")

# 공유 legend
legend_shared <- get_legend(
  gg_drought1_Agr11 + theme(legend.position = "bottom")
)

plot_grid(main_grid, legend_shared, ncol = 1, rel_heights = c(10, 1))


ggsave("figures/fig_drought1.png")



```


## Drought Event 2 (2014–2017) 

```{r, fig.width=10, fig.height=5}
p1 <- gg_drought2_Agr5 + 
  ggtitle("a)") + 
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0, face = "bold", size = 14))

p2 <- gg_drought2_Agr7 + 
  ggtitle("b)") + 
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0, face = "bold", size = 14),
        axis.title.y = element_blank())  # ← 오른쪽 plot y축 레이블 제거

main_grid <- plot_grid(p1, p2, ncol = 2, align = "hv", axis = "tblr")

# 공유 legend
legend_shared <- get_legend(
  gg_drought1_Agr11 + theme(legend.position = "bottom")
)

plot_grid(main_grid, legend_shared, ncol = 1, rel_heights = c(10, 1))

ggsave("figures/fig_drought2.png")

```


# Cumulative lines

## recipients 

```{r}
cumline <- data.frame(days = 1:c(39*365), cumsum_withoutABM= cumsum(without_ABM_new_diversion$df_agr[,5]),cumsum_withABM= cumsum(with_ABM_new_diversion$df_agr[,5]))
r1 <- ggplot(data = cumline, aes(x = days)) +
  geom_line(aes(y = cumsum_withoutABM, color = "Without ABM")) +
  geom_line(aes(y = cumsum_withABM, color = "With ABM")) +
  scale_color_manual(
    values = c("Without ABM" = "blue", "With ABM" = "red"),
    name = "Scenario") +
  labs(
    title = "Agr 5",
    x = "Days",
    y = "Cumulative withdrawal (MCM)"
  ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))

r1
```

```{r}
## Intake 11
cumline <- data.frame(days = 1:c(39*365), cumsum_withoutABM= cumsum(without_ABM_new_diversion$df_intake[,11]),cumsum_withABM= cumsum(with_ABM_new_diversion$df_intake[,11]))
d3 <- ggplot(data = cumline, aes(x = days)) +
  geom_line(aes(y = cumsum_withoutABM, color = "Without ABM")) +
  geom_line(aes(y = cumsum_withABM, color = "With ABM")) +
  scale_color_manual(
    values = c("Without ABM" = "blue", "With ABM" = "red"),
   name = "Scenario" ) +
  labs(
    title = "Intake 11",
    x = "Days",
    y = "Cumulative withdrawal (MCM)"
  ) +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))

d3
```

```{r}

## Agr 12 recipient
cumline <- data.frame(days = 1:c(39*365), cumsum_withoutABM= cumsum(without_ABM_new_diversion$df_agr[,13]),cumsum_withABM= cumsum(with_ABM_new_diversion$df_agr[,13]))
r4 <- ggplot(data = cumline, aes(x = days)) +
  geom_line(aes(y = cumsum_withoutABM, color = "Without ABM")) +
  geom_line(aes(y = cumsum_withABM, color = "With ABM")) +
  scale_color_manual(
    values = c("Without ABM" = "blue", "With ABM" = "red"),
    name = "Scenario" ) +
  labs(
    title = "Agr 13",
    x = "Days",
    y = "Cumulative withdrawal (MCM)"
  ) +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))

r4

```

```{r}
# Agr 7
cumline <- data.frame(days = 1:c(39*365), cumsum_withoutABM= cumsum(without_ABM_new_diversion$df_agr[,7]),cumsum_withABM= cumsum(with_ABM_new_diversion$df_agr[,7]))
r2<-ggplot(data = cumline, aes(x = days)) +
  geom_line(aes(y = cumsum_withoutABM, color = "Without ABM")) +
  geom_line(aes(y = cumsum_withABM, color = "With ABM")) +
   scale_color_manual(
    values = c("Without ABM" = "blue", "With ABM" = "red"),
    name = "Scenario") +
  labs(
    title = "Agr 7",
    x = "Days",
    y = "Cumulative withdrawal (MCM)"
  ) +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))

r2

```

```{r}


# Agr 10
cumline <- data.frame(days = 1:c(39*365), cumsum_withoutABM= cumsum(without_ABM_new_diversion$df_agr[,11]),cumsum_withABM= cumsum(with_ABM_new_diversion$df_agr[,11]))
r3<-ggplot(data = cumline, aes(x = days)) +
  geom_line(aes(y = cumsum_withoutABM, color = "Without ABM")) +
  geom_line(aes(y = cumsum_withABM, color = "With ABM")) +
   scale_color_manual(
    values = c("Without ABM" = "blue", "With ABM" = "red"),
    name = "Scenario") +
  labs(
    title = "Agr 11",
    x = "Days",
    y = "Cumulative withdrawal (MCM)"
  ) +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))
r3
```

```{r}

## donor
cumline <- data.frame(days = 1:c(39*365), cumsum_withoutABM= cumsum(without_ABM_new_diversion$df_agr[,13]),cumsum_withABM= cumsum(with_ABM_new_diversion$df_agr[,13]))
r4<-ggplot(data = cumline, aes(x = days)) +
  geom_line(aes(y = cumsum_withoutABM, color = "Without ABM")) +
  geom_line(aes(y = cumsum_withABM, color = "With ABM")) +
  scale_color_manual(
    values = c("Without ABM" = "blue", "With ABM" = "red"),
    name = "Scenario" ) +
  labs(
    title = "Agr 13",
    x = "Days",
    y = "Cumulative withdrawal (MCM)"
  ) +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))

r4

```

## donors

```{r}
cumline <- data.frame(days = 1:c(39*365), cumsum_withoutABM= cumsum(without_ABM_new_diversion$df_intake[,4]),cumsum_withABM= cumsum(with_ABM_new_diversion$df_intake[,4]))
d1<-ggplot(data = cumline, aes(x = days)) +
  geom_line(aes(y = cumsum_withoutABM, color = "Without ABM")) +
  geom_line(aes(y = cumsum_withABM, color = "With ABM")) +
  scale_color_manual(
    values = c("Without ABM" = "blue", "With ABM" = "red"),
    name = "Scenario" ) +
  labs(
    title = "Intake 4",
    x = "Days",
    y = "Cumulative withdrawal (MCM)"
  ) +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))

d1
```

```{r}
cumline <- data.frame(days = 1:c(39*365), cumsum_withoutABM= cumsum(without_ABM_new_diversion$df_agr[,12]),cumsum_withABM= cumsum(with_ABM_new_diversion$df_agr[,12]))
ggplot(data = cumline, aes(x = days)) +
  geom_line(aes(y = cumsum_withoutABM, color = "Without ABM")) +
  geom_line(aes(y = cumsum_withABM, color = "With ABM")) +
  #scale_color_manual(
  #  values = c("Without ABM" = "blue", "With ABM" = "red"),
  #  name = "Scenario" ) +
  labs(
    title = "Agr 12",
    x = "Days",
    y = "Cumulative withdrawal (MCM)"
  ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
cumline <- data.frame(days = 1:c(39*365), cumsum_withoutABM= cumsum(without_ABM_new_diversion$df_intake[,11]),cumsum_withABM= cumsum(with_ABM_new_diversion$df_intake[,11]))
d3<-ggplot(data = cumline, aes(x = days)) +
  geom_line(aes(y = cumsum_withoutABM, color = "Without ABM")) +
  geom_line(aes(y = cumsum_withABM, color = "With ABM")) +
  scale_color_manual(
    values = c("Without ABM" = "blue", "With ABM" = "red"),
    name = "Scenario") +
  labs(
    title = "Intake 11",
    x = "Days",
    y = "Cumulative withdrawal (MCM)"
  ) +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))
d3
```

```{r}

## donor
cumline <- data.frame(days = 1:c(39*365), cumsum_withoutABM= cumsum(without_ABM_new_diversion$df_agr[,14]),cumsum_withABM= cumsum(with_ABM_new_diversion$df_agr[,14]))
d4<-ggplot(data = cumline, aes(x = days)) +
  geom_line(aes(y = cumsum_withoutABM, color = "Without ABM")) +
  geom_line(aes(y = cumsum_withABM, color = "With ABM")) +
  scale_color_manual(
    values = c("Without ABM" = "blue", "With ABM" = "red"),
    name = "Scenario" ) +
  labs(
    title = "Agr 14",
    x = "Days",
    y = "Cumulative withdrawal (MCM)"
  ) +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))

d4
```

```{r}
cumline <- data.frame(days = 1:c(39*365), cumsum_withoutABM= cumsum(without_ABM_new_diversion$df_intake[,6]),cumsum_withABM= cumsum(with_ABM_new_diversion$df_intake[,6]))
d2<-ggplot(data = cumline, aes(x = days)) +
  geom_line(aes(y = cumsum_withoutABM, color = "Without ABM")) +
  geom_line(aes(y = cumsum_withABM, color = "With ABM")) +
  scale_color_manual(
    values = c("Without ABM" = "blue", "With ABM" = "red"),
    name = "Scenario") +
  labs(
    title = "Intake 6",
    x = "Days",
    y = "Cumulative withdrawal (MCM)"
  ) +
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))
d2
```

```{r, fig.width = 8, fig.height = 12}
library(gridExtra)
library(cowplot) 
cumline <- data.frame(
  days = 1:c(39*365),
  cumsum_withoutABM = cumsum(without_ABM_new_diversion$df_intake[,6]),
  cumsum_withABM = cumsum(with_ABM_new_diversion$df_intake[,6])
)

# ✅ 공유 legend용 플롯
p_legend <- ggplot(data = cumline, aes(x = days)) +
  geom_line(aes(y = cumsum_withoutABM, color = "Without ABM")) +
  geom_line(aes(y = cumsum_withABM, color = "With ABM")) +
  scale_color_manual(
    values = c("Without ABM" = "blue", "With ABM" = "red"),
    name = "Scenario"
  ) +
  theme_classic() +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 16, face = "bold"),
    legend.text  = element_text(size = 16
                                )
  )
r1 <- r1 + ggtitle("a) Agr 5") + theme(legend.position = "none", plot.title = element_text(hjust = 0, face = "bold", size = 16))
d1 <- d1 + ggtitle("b) Intake 4") + theme(legend.position = "none", plot.title = element_text(hjust = 0, face = "bold", size = 16))
r2 <- r2 + ggtitle("c) Agr 7") + theme(legend.position = "none", plot.title = element_text(hjust = 0, face = "bold", size = 16))
d2 <- d2 + ggtitle("d) Intake 6") + theme(legend.position = "none", plot.title = element_text(hjust = 0, face = "bold", size = 16))
r3 <- r3 + ggtitle("e) Agr 11") + theme(legend.position = "none", plot.title = element_text(hjust = 0, face = "bold", size = 16))
d3 <- d3 + ggtitle("f) Intake 11") + theme(legend.position = "none", plot.title = element_text(hjust = 0, face = "bold", size = 16))
r4 <- r4 + ggtitle("g) Agr 13") + theme(legend.position = "none", plot.title = element_text(hjust = 0, face = "bold", size = 16))
d4 <- d4 + ggtitle("h) Agr 14") + theme(legend.position = "none", plot.title = element_text(hjust = 0, face = "bold", size = 16))

# ✅ 2. 크기 정렬
aligned <- align_plots(r1, d1, r2, d2,
                       r3, d3, r4, d4,
                       align = "hv",
                       axis = "tblr")

# ✅ 3. 그리드 배치
main_grid <- plot_grid(
  aligned[[1]], aligned[[2]],
  aligned[[3]], aligned[[4]],
  aligned[[5]], aligned[[6]],
  aligned[[7]], aligned[[8]],
  ncol = 2, nrow = 4
)

# ✅ 4. legend 추출
legend_grob <- get_legend(p_legend)

# ✅ 5. 최종 출력
plot_grid(
  main_grid,
  legend_grob,
  ncol = 1,
  rel_heights = c(20, 1)
)

ggsave("figures/fig8.png", width = 12, height = 16, dpi = 300)
```
```{r}
with_ABM_new_diversion$breaker
```




```{r validation}

df_ratio <- as.data.frame(without_ABM_new_diversion$df_agr / without_ABM_new_diversion$demand_agr)
colnames(df_ratio) <- paste0("Agr", seq_len(ncol(df_ratio)))

start_date <- as.Date("1981-10-01")
df_ratio <- df_ratio %>%
  mutate(Date = seq(start_date, by = "day", length.out = nrow(df_ratio))) %>%
  relocate(Date)

df_2018 <- df_ratio %>%
  filter(Date >= as.Date("2018-01-01"),
         Date <= as.Date("2018-12-31"))

mat_2018 <- as.matrix(select(df_2018, -Date))
mat_2018[!is.finite(mat_2018)] <- NA  # Inf, -Inf, NaN -> NA

lack_days_2018 <- 153 - colSums(mat_2018, na.rm = TRUE)



df_2019 <- df_ratio %>%
  filter(Date >= as.Date("2019-01-01"),
         Date <= as.Date("2019-12-31"))

mat_2019 <- as.matrix(select(df_2019, -Date))
mat_2019[!is.finite(mat_2019)] <- NA  # Inf, -Inf, NaN -> NA

lack_days_2019 <- 153 - colSums(mat_2019, na.rm = TRUE)


df_2020 <- df_ratio %>%
  filter(Date >= as.Date("2020-01-01"),
         Date <= as.Date("2020-12-31"))

mat_2020 <- as.matrix(select(df_2020, -Date))
mat_2020[!is.finite(mat_2020)] <- NA  # Inf, -Inf, NaN -> NA

lack_days_2020  <- 153 - colSums(mat_2020, na.rm = TRUE)
lack_days_2018
lack_days_2019
lack_days_2020


```
```{r}
# ── 1. Reliability ratio ───────────────────────────────────────────────────
df_ratio <- as.data.frame(
  without_ABM_new_diversion$df_agr / without_ABM_new_diversion$demand_agr)
colnames(df_ratio) <- paste0("Agr", seq_len(ncol(df_ratio)))

start_date <- as.Date("1981-10-01")
df_ratio <- df_ratio %>%
  mutate(Date = seq(start_date, by = "day", length.out = nrow(df_ratio))) %>%
  relocate(Date)

# ── 2. Water shortage days 계산 (관개기 153일 기준) ────────────────────────
calc_lack <- function(df_ratio, year) {
  df_yr  <- df_ratio %>%
    filter(Date >= as.Date(paste0(year, "-01-01")),
           Date <= as.Date(paste0(year, "-12-31")))
  mat <- as.matrix(select(df_yr, -Date))
  mat[!is.finite(mat)] <- NA
  # 공급 부족일: ratio < 1인 날 합산 (153 - sum(ratio)보다 정확)
  shortage <- colSums(mat < 1, na.rm = TRUE)
  shortage
}

lack_days_2018 <- calc_lack(df_ratio, 2018)
lack_days_2019 <- calc_lack(df_ratio, 2019)
lack_days_2020 <- calc_lack(df_ratio, 2020)

lack_days_2018
lack_days_2019
lack_days_2020

# ── 3. 실측 논 피해면적 (Table 1, National Drought Info Statistics 2020) ──
paddy_wilting_ha <- c(
  Agr1  = 0.0000,
  Agr2  = 0.0000,
  Agr3  = 0.5653,
  Agr4  = 3.1867,
  Agr5  = 19.3849,
  Agr6  = 1.1143,
  Agr7  = 5.1818,
  Agr8  = 3.9110,
  Agr9  = 4.7674,
  Agr10 = 0.7286,
  Agr11 = 47.6349,
  Agr12 = 62.7743,
  Agr13 = 34.5280,
  Agr14 = 42.0238
)

# ── 4. Spearman 상관 (2018 기준) ──────────────────────────────────────────
validation_df <- data.frame(
  Subbasin     = names(lack_days_2018),
  shortage_days = as.numeric(lack_days_2018),
  paddy_ha      = paddy_wilting_ha
)

# rank
validation_df <- validation_df %>%
  mutate(
    rank_shortage = rank(-shortage_days, ties.method = "min"),
    rank_paddy    = rank(-paddy_ha,      ties.method = "min")
  )

spearman_result <- cor.test(
  validation_df$shortage_days,
  validation_df$paddy_ha,
  method = "spearman"
)

cat("=== Spearman Rank Correlation (2018) ===\n")
cat(sprintf("rho = %.3f\n", spearman_result$estimate))
cat(sprintf("p   = %.3f\n", spearman_result$p.value))
cat(sprintf("N   = %d\n",   nrow(validation_df)))

table1 <- validation_df %>%
  select(Subbasin, shortage_days, rank_shortage, 
         paddy_ha, rank_paddy)

colnames(table1) <- c(
  "Subbasin",
  "Water shortage days",
  "Rank (shortage)",
  "Paddy wilting (ha)",
  "Rank (paddy)"
)

print(table1)
```


```{r table7}

  # diversion 1: Node B --> Agr5
  # diversion 2: Node C --> Agr7
  # diversion 3: Node E --> Agr11
  # diversion 4: Node F --> Agr13
  # diversion 5: Intake 4 --> Agr5
  # diversion 6: Intake 6 --> Agr7
  # diversion 7: Intake 11 --> Agr11

start_date <- as.Date("1981-10-01")
n_days <- nrow(with_ABM_new_diversion$help)

date_seq <- seq(start_date, by = "day", length.out = n_days)

result <- data.frame(Date = date_seq, help= with_ABM_new_diversion$help)

result


result_E1 <- result |>
  filter(
    Date >= as.Date("1994-01-01"),
    Date <= as.Date("1997-12-31")
  )
result_E2 <- result |>
  filter(
    Date >= as.Date("2014-01-01"),
    Date <= as.Date("2017-12-31")
  )
print("Event 1")
colSums(result_E1[2:8])
sum(colSums(result_E1[2:8]))
print("Event 2")
colSums(result_E2[2:8])
sum(colSums(result_E2[2:8]))



sum(colSums(result_E2[2:8])) - sum(colSums(result_E1[2:8]))
```

```{r}


activation <- data.frame(Date = date_seq, help= with_ABM_new_diversion$activation)

activation_E1 <- activation |>
  filter(
    Date >= as.Date("1994-01-01"),
    Date <= as.Date("1997-12-31")
  )

activation_E2 <- activation |>
  filter(
    Date >= as.Date("2014-01-01"),
    Date <= as.Date("2017-12-31")
  )
print("Event 1")
colSums(activation_E1[2:8])
sum(colSums(activation_E1[2:8]))
print("Event 2")
colSums(activation_E2[2:8])
sum(colSums(activation_E2[2:8]))

```

```{r, width=120}

rel_with <- data.frame(Date = date_seq,
                  With = with_ABM_new_diversion$df_agr /with_ABM_new_diversion$demand_agr)

rel_without <- data.frame(Date = date_seq,
                  Without = without_ABM_new_diversion$df_agr /without_ABM_new_diversion$demand_agr)

rel_with_E1 <- rel_with |>
  filter(
    Date >= as.Date("1994-01-01"),
    Date <= as.Date("1997-12-31")
  )

rel_without_E1 <- rel_without |>
  filter(
    Date >= as.Date("1994-01-01"),
    Date <= as.Date("1997-12-31")
  )

rel_with_E2 <- rel_with |>
  filter(
    Date >= as.Date("2014-01-01"),
    Date <= as.Date("2017-12-31")
  )

rel_without_E2 <- rel_without |>
  filter(
    Date >= as.Date("2014-01-01"),
    Date <= as.Date("2017-12-31")
  )

print("Event 1: with")
print(colSums(rel_with_E1[-1],na.rm=TRUE ))
#sum(colSums(rel_with_E1[2:8],na.rm=TRUE ))

print("Event 1: without")
print(colSums(rel_without_E1[-1],na.rm=TRUE ))
#sum(colSums(rel_without_E1[2:8],na.rm=TRUE ))

print("Event 2: with")
print(colSums(rel_with_E2[-1],na.rm=TRUE ))
#sum(colSums(rel_with_E2[2:8],na.rm=TRUE ))

print("Event 2: without")
print(colSums(rel_without_E2[-1],na.rm=TRUE ))
#sum(colSums(rel_without_E2[2:8],na.rm=TRUE ))


```






